
\chapter*{Acknowledgements}

Finally, the time has come to write acknowledgements, which makes me both sad and relieved. Luckily, writing these lines I also think back to all the fun I had, which I almost forgot after the tiring task of writing this book. This book, which I thought nobody might read apart from the people who helped me write it and improve it. However, it turns out people do read it, and I would like to thank the members of my jury for reading the manuscript with so much interest and in so much detail, for the interesting questions and discussions, and for the valuable comments.

More importantly, I would like to thank my promotor, Steven Lowette, for giving me the opportunity to start this whole adventure in the first place, more than 4 years ago. I am very grateful to you for guiding me through these 2 dark matter searches and giving me the freedom to combine this with various projects I pursued in CMS. I enjoyed working on the SIMP pheno study and CMS analysis a lot, and I would like to thank Michel and Bryan for the very nice collaboration and for discussing and answering our many questions which keep coming, even at the very end of this PhD.

I started my PhD by getting to know the detector a bit better and went on an incredible tracker-adventure together with Isis and Lieselotte. Thank you for the all the fun we had in Brussels, but also the great times we had travelling to CERN, staying at the apartment, and working at P5. We are a great team and learned a lot from each other, about physics, the tracker, cooking, and many more things. I wish you both the best of luck with your own thesis, you are almost there!

During my stays at CERN I got to know many extraordinary people, probably too many to list here. I met nice colleagues I worked with on the monojet analysis, and I got to know the long-lived working group, where it was always nice to present my work. Going to CERN, I spent most of my time at P5 doing shifts and getting to know the tracker better, and I met a lot of people I would like to thank for showing me the way at P5 on the surface, underground, and even inside the detector. Taking the risk of forgetting someone, I would like to thank some of the tracker people I worked with: Derek, Franz, Christian, Ferdinando, Nicola, and Stefanos. I will never forget your enthusiasm whenever we brought cake. I would also like to especially thank Erik, the scary ops manager and Hero of the largest silicon tracker in the world, for all the advice, the encouragements, and for reading all this text and subsequently crushing my soul with many comments and suggestions. You helped me to keep going and greatly improved the quality of my manuscript. Becoming co-convener of the tracker DQM group, I got to know even more people from the tracker group and I would like to thank them, and specifically my co-conveners Hugo and Francesco.

Over the past 4 years, the people I shared this huge office with have been coming and going, but the pleasant atmosphere never left. Hoping I do not forget anyone, I would like to thank James, Annik, Gerrit, Kevin, Denys, Shimaa, Seth, Kirill, Leonidas, Petra, Doug, Simon, Dominic, Emil, and -- even though he is a traitor who abandoned our office -- Jarne. I also want to thank 2 more people from the office, Nadir and Giannis, with whom I had the pleasure of working on the SIMP-undertaking, and who helped and taught me a lot. I would like to thank all my other wonderful colleagues at the IIHE for the great time I spent here, filled with awesome coffee breaks and amazing pie. Discovering many of you also play music, I enjoyed all the rehearsals with the IIHE band. I would also like to thank the admins and technical support, particularly for saving my laptop twice, the second time even bringing it back from the dead! Finally, I want to thank Marleen, without her organisation and knowledge we would not survive (or have nice receptions and parties).

Additionally, I want to thank my family, especially my parents, sister, and parents-in-law, for all the help, company, and support during the difficult and lonely moments of my PhD. Special thanks to my sister, Suzanne, for designing the Strips logo and a very nice cover for this book. I also want to thank my friends, for staying in contact during this busy period and for organising dinners, get-togethers, and a weekend break. Particularly, I want to thank Sander for helping so much in the house. I wish you a lot of perseverance to finish your thesis.

Last but not least, I want to thank my husband Hendrik for always encouraging me to be ambitious and to pursue my goals. Thank you for letting me leave you time and again to go to CERN. We made some ambitious decisions and some people might say we are crazy, but we got through this difficult period. Even though you fled to Switzerland, I didn't let you escape from all my complaints and frustration, and continued to harass you over the phone. Thank you for the unconditional support and encouragements during these 4 years and especially during these trying times also for you.